---
title: "Inflation Data"
author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
repro:
  packages:
    - here
    - tidyverse
    - lubridate
    - aaronpeikert/repro@fc7e884
    - papaja
  scripts:
    - R/prepare_inflation.R
  data:
    - data/raw/inflation.rds
---

```{r setup, include=FALSE}
repro::automate_load_packages()
repro::automate_load_scripts()
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library("papaja")
r_refs("r-references.bib")
```

The dataset we use stems from the [Bank of England Research datasets](https://www.bankofengland.co.uk/statistics/research-datasets).

I quote:

> This dataset contains the individual responses to our Inflation Attitudes Survey, a quarterly survey of people’s feelings about inflation and other economic variables like the interest rate. 

```{r load-inflation}
# code chunks can come from other scripts, this one is loaded from:
# `R/prepare_inflation.R`
```

```{r transform-inflation}
# this chunk is also loaded from the same script
# note that they must be empty and the names must correspond
```

For this dataset the Bank of England asked `r prettyNum(nrow(inflation), big.mark = ".", decimal.mark = ",")` people for their opinion on the perceived and expected inflation.
The survey has run quarterly since `r min(inflation$year)`.

```{r}
inflation %>% 
  group_by(date) %>% 
  summarise(across(c(perception, expectation),
                   ~ mean(., na.rm = TRUE)),
            .groups = "drop") %>% 
  pivot_longer(c(expectation, perception)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_line(aes(date, value, color = name)) +
  NULL
```

Inflation is an economic phenomenon characterized by a rise in the general price level of goods and services within an economy over a period of time. 
When inflation occurs, each unit of currency buys fewer goods and services, indicating a loss of purchasing power. 
It's usually measured by the Consumer Price Index (CPI) or the Producer Price Index (PPI). 
Inflation can be triggered by various factors including increased production costs, higher demand for goods and services, and monetary policies. 
Central banks often aim to control inflation to ensure it remains at a manageable level, as moderate inflation is seen as a sign of a growing economy, while hyperinflation or deflation can be detrimental.
